*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-29 Code/Manuscript_O2.Rmd
minor edits

Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. Photosynthetically Usable Radiation (PUR) represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll *a* and 478 nm for divinyl chlorophyll *b*
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth. We therefore analyzed growth rates in terms of cumulative diel PUR.

We discuss our findings in relation to analyses of genomic sequences CITATIONS across clades of *P. marinus*, showing that differences in the presence of genes encoding oxygen-dependent enzymes, and DNA repair enzymes, can help explain differential growth responses of strains under the matrix of light and [O~2~] conditions.

@@

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focussed on proteins mediating photosynthesis and protein metabolism. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).

Collection and treatment of protein samples were performed by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

>>>>

Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. Photosynthetically Usable Radiation (PUR) represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes show an absorption maxima
of 442 nm for divinyl chlorophyll *a* and 478 nm for divinyl chlorophyll *b* [@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to efficiently harvest blue light in the 400 nm to 500 nm range [@morelAvailableUsableStored1978] where only blue spectral wavelengths prevail in deep ocean habitats [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993] contributing to efficient optical absorption. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth. We therefore analyzed growth rates in terms of cumulative diel PUR.

We discuss our findings in relation to analyses of genomic sequences [@changBRENDAELIXIRCore2021] across clades of *P. marinus*, showing that differences in the presence of genes encoding oxygen-dependent enzymes, and DNA repair enzymes, can help explain differential growth responses of strains under the matrix of light and [O~2~] conditions of this study.

@@

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focused on proteins mediating photosynthesis and protein metabolism. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).

Collection and treatment of protein samples were performed by personnel at the WHOI according to the protocols described by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.
